---
title: "Decompose Variance"
author: "Sarah Teichman"
date: "04/30/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bayesplot)
library(rstan)
library(tidyverse)
theme_update(plot.title = element_text(hjust = 0.5))
load("fit_ifls_4-30-21.rda")
load(file = "data_no_s6.rda")
```

```{r, echo = F}
dat_df <- data_no_s6
K <- dat_df$K
Ti <- dat_df$T
N <- dat_df$N
log_y <- log(dat_df$y)
job <- array(data = NA, dim = c(N,Ti))
for (i in 1:N) {
  for (t in 1:Ti) {
    job[i,t] <- which(dat_df$a[i,t,]==1)
  }
}
dat <- data.frame(log_wage = c(log_y[,1],log_y[,2],log_y[,3]),
                  worker = rep(1:N, 3),
                  sector = c(job[,1], job[,2], job[,3]))
```

## Compare worker and sector effects between AKM and our model 

```{r,echo = F}
fit <- fit_ifls
post <- as.matrix(fit)
fit_summ <- as.data.frame(summary(fit)$summary)
mu_params <- grep("mu", row.names(fit_summ), value = T)
nu_params <- grep("nu", row.names(fit_summ), value = T)
f_params <- grep("unique_f", row.names(fit_summ), value = T)
```

### AKM variance results 

```{r, echo = F}
tot_var <- var(dat$log_wage)
akm_df <- data.frame(log_wage = c(log_y[,1],log_y[,2],log_y[,3]),
                  worker = as.factor(rep(1:N,Ti)),
                  sector = as.factor(c(job[,1],job[,2],job[,3])))
mod <- lm(log_wage ~ -1 + worker + sector, data = akm_df)
akm_worker_eff <- mod$coefficients[1:N]
akm_sector_eff <- mod$coefficients[N:length(mod$coefficients)]
akm_sector_eff[1] <- 0 
worker_var <- var(akm_worker_eff)/tot_var
sector_var <- var(akm_sector_eff)/tot_var
df <- akm_df %>%
  mutate(worker_effect = akm_worker_eff[worker],
         sector_effect = akm_sector_eff[sector])
worker_sector_cov <- cov(df$worker_effect, df$sector_effect)
worker_var # 56% of variance from worker
sector_var # 5% of variance from worker
2*worker_sector_cov/tot_var # 3% of variance from covariance between worker and sector 
```

### Comparison between AKM effects and our model effect means

```{r, echo = F}
worker_df <- data.frame(akm_worker = akm_worker_eff,
                        mod_worker = fit_summ[mu_params, 1])
ggplot(worker_df, aes(x = akm_worker, y = mod_worker)) +
  geom_point() + 
  xlab("AKM worker effect") + 
  ylab("Model worker effect")
sector_df <- data.frame(akm_sector = akm_sector_eff,
                        mod_sector = fit_summ[nu_params, 1])
ggplot(sector_df, aes(x = akm_sector, y = mod_sector)) +
  geom_point() + 
  xlab("AKM sector effect") + 
  ylab("Model sector effect")
cor(worker_df$akm_worker, worker_df$mod_worker)
cor(sector_df$akm_sector, sector_df$mod_sector)
```

## Variance of posterior means in our model

```{r}
#worker_var <- var(dat$mu)
worker_var <- var(fit_summ[mu_params,1])
#sector_var <- var(dat$nu)
sector_var <- var(fit_summ[nu_params,1])
firm_var <- var(fit_summ[f_params,1])
tot_var
worker_var/tot_var
sector_var/tot_var
firm_var/tot_var
```

## Mean of posterior variances in our model 

```{r,echo = F}
post_mu <- post[,mu_params]
post_nu <- post[,nu_params]
post_f <- post[,f_params]
mu_var <- apply(post_mu, 1, var)
nu_var <- apply(post_nu, 1, var)
f_var <- apply(post_f, 1, var)
mu_mean_var <- mean(mu_var)
nu_mean_var <- mean(nu_var)
f_mean_var <- mean(f_var)
```

```{r}
tot_var <- var(dat$log_wage)
mu_mean_var / tot_var
nu_mean_var / tot_var 
f_mean_var / tot_var
```

```{r, echo = F}
plot_df <- data.frame(mu_var = mu_var/tot_var, nu_var = nu_var/tot_var,
                      f_var = f_var/tot_var) 
ggplot(plot_df, aes(x = mu_var)) + geom_histogram() +
  xlab("Proportion variance in worker effects") + ylab("Count")
ggplot(plot_df, aes(x = nu_var)) + geom_histogram() +
  xlab("Proportion variance in sector effects") + ylab("Counts")
ggplot(plot_df, aes(x = f_var)) + geom_histogram() +
  xlab("Proportion variance in firm effects") + ylab("Counts")
```

## Mean of Posterior Covariances

```{r,echo = F}
covar <- vector(length = nrow(post))
for (i in 1:nrow(post)) {
  worker <- rep(post[i,mu_params], 3)
  firm <- post[i, c(f_params[dat_df$firm_effect_ind[,1]],
                      f_params[dat_df$firm_effect_ind[,2]],
                      f_params[dat_df$firm_effect_ind[,3]])]
  worker <- unlist(worker)
  firm <- unlist(firm)
  covar[i] <- cov(worker, firm)
}
mean(covar)
cov_df <- data.frame(covar = 2*covar/tot_var) 
ggplot(cov_df, aes(x = covar)) + geom_histogram() +
  xlab("Proportion 2*covariance in worker and firm effects") + ylab("Count")
```

## Calculate correlation between AKM and our model worker and sector effects for each iteration

```{r}
corr_worker <- vector(length = nrow(post))
corr_sector <- vector(length = nrow(post))
for (i in 1:nrow(post)) {
  corr_worker[i] <- cor(post[i,mu_params], akm_worker_eff)
  corr_sector[i] <- cor(post[i,nu_params], akm_sector_eff)
}
mean(corr_worker)
mean(corr_sector)
```


## Overall wage decompostion 

$$Var(f_j) = E[Var(f_j|j\in k)] + Var(E[f_j|j\in k])$$

The second term is estimated by `r nu_mean_var` (nu_mean_var). The second term can be estimated by taking the average over the $\frac{1}{\tau_k}$ for each sector $k$. 

We will consider the mean over the posterior mean of variance terms $\frac{1}{\tau_k}$ in each iteration. 

```{r, echo = FALSE}
tau_k_params <- grep("tau_k", row.names(fit_summ), value = T)
post_tau_k <- post[,tau_k_params]
var_k_means <- apply(1/post_tau_k, 1, mean)
var_k_mean <- mean(var_k_means)

(var_k_mean + nu_mean_var)
(var_k_mean + nu_mean_var)/tot_var

```

Finally, in each iteration I'll get the proportion variance associated with worker and with sector. 

```{r, echo = FALSE}
prop_var <- matrix(data = NA, nrow(post), 4) 
for (i in 1:nrow(post)) {
  worker_var <- var(post[i, mu_params])
  nu_var <- var(post[i, nu_params])
  k_mean_var <- mean(1/post[i, tau_k_params])
  prop_var[i,1] <- worker_var/tot_var
  prop_var[i,2] <- nu_var/tot_var
  prop_var[i,3] <- k_mean_var/tot_var
  prop_var[i,4] <- (nu_var+k_mean_var)/tot_var
}

summary(prop_var)
hist(prop_var[,1])
hist(prop_var[,4])
```