---
title: "Parameter Interpretation - IFLS data"
author: "Sarah Teichman"
date: "04/13/20"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bayesplot)
library(rstan)
library(tidyverse)
library(splitstackshape)
library(stringi)
library(parallelDist)
library(labdsv)
library(ape)
load("fit_ifls_4-11.rda")
load("deflate_y_dat.rda")
```

```{r,echo=F}
fit <- fit_ifls
dat <- deflate_y_dat
post <- as.data.frame(fit)
fit_summ <- as.data.frame(summary(fit)$summary)
```

```{r,echo=F}
K <- 7
Ti <- 3
N <- 1973
```

#### Posterior histograms for $\sigma^2_k$ 

Plotting posterior histograms from $\sigma^2_k$ parameters. 

```{r, echo = F}
n_sim <- length(post$sigmasq_w0)
labs <- c("Agriculture","Manufacturing","Construction","Hospitality",
          "Transportation","Business","Social Services")
post_hist <- data.frame(val = c(post$`sigmasq_k[1]`,post$`sigmasq_k[2]`,post$`sigmasq_k[3]`,
                                post$`sigmasq_k[4]`,post$`sigmasq_k[5]`,post$`sigmasq_k[6]`,
                                post$`sigmasq_k[7]`),
                        class = c(rep("class1",n_sim),rep("class2",n_sim),rep("class3",n_sim),
                                 rep("class4",n_sim),rep("class5",n_sim),rep("class6",n_sim),
                                 rep("class7",n_sim))) %>%
  mutate(class = labs[class])
```

Computing empirical wage variances across firm classes. 

```{r, echo = F}
log_y <- log(dat$y)
job <- array(data = NA, dim = c(N,Ti))
for (i in 1:N) {
  for (t in 1:Ti) {
    job[i,t] <- which(dat$a[i,t,]==1)
  }
}
wage_job_dat <- data.frame(worker = rep(1:N,3), time = sort(rep(1:3,N)),
                            log_wage = c(log_y[,1],log_y[,2],log_y[,3]),
                            job = c(job[,1],job[,2],job[,3])) %>%
  mutate(job = labs[job])
#wage_job_dat$job <- as.factor(wage_job_dat$job)
class_var <- wage_job_dat %>%
  group_by(job) %>%
  summarise(variance = var(log_wage)) %>% 
  mutate(class = job)
ggplot() + geom_histogram(data = post_hist, aes(x = val)) + 
  geom_vline(data = class_var, aes(xintercept = variance), color = "red") + 
  facet_wrap(~class) + xlab("variance") + ggtitle("Posterior Variances for Classes")
```

#### Size of each Class over Time 

```{r,echo=F}
job_long <- as.data.frame(job) %>% pivot_longer(cols = c(V1,V2,V3),names_to = "time",
                                                names_prefix = "V") %>%
  mutate(class = labs[value])
order <- job_long %>% group_by(class, time) %>%
  summarise(count = n()) %>%
  arrange(time, count)
job_long$class <- factor(job_long$class, levels = order$class[1:7])
ggplot(job_long, aes(x = class)) + geom_bar() + facet_wrap(~time) + 
  xlab("Class") + theme(axis.text.x = element_text(angle = 270, hjust = 1))
```

#### Log wages by Class and Time 

```{r,echo=F}
meds <- wage_job_dat %>% group_by(time, job) %>%
  summarise(med = median(log_wage)) %>%
  arrange(time, med)
wage_job_dat$job <- factor(wage_job_dat$job, levels = meds$job[1:7])
ggplot(wage_job_dat, aes(x = job, y = log_wage)) + geom_boxplot() +
  theme(axis.text.x = element_text(angle = 270, hjust = 1)) +
  facet_wrap(~time) + xlab("Sector") + ylab("Log Wage")
```

#### Plot latent space positions 

```{r,echo=F}
med_z <- data.frame(par = rownames(fit_summ), med = fit_summ[,6]) %>%
  filter(grepl("z\\[", par)) %>%
  cSplit("par",direction = "wide") %>%
  mutate(worker = stri_extract_first_regex(par_1,"[0-9]+"),
         dim = stri_extract_first_regex(par_2,"[0-9]+")) %>%
  select(-c("par_1","par_2")) %>%
  pivot_wider(names_from = dim, values_from = med) %>%
  mutate(job1 = job[,1],
         job2 = job[,2],
         job3 = job[,3]) %>%
  pivot_longer(cols = c(job1,job2,job3),names_to = "time",names_prefix = "job") %>% 
  mutate(class = labs[value]) %>%
  select(-value) 
med_w <- data.frame(par = rownames(fit_summ), med = fit_summ[,6]) %>%
  filter(grepl("w\\[", par)) %>%
  cSplit("par",direction = "wide") %>%
  mutate(class = stri_extract_first_regex(par_1,"[0-9]+"),
         time = stri_extract_first_regex(par_2,"[0-9]+"),
         dim = stri_extract_first_regex(par_3,"[0-9]+")) %>%
  select(-c("par_1","par_2","par_3"))%>%
  pivot_wider(names_from = dim, values_from = med) %>%
  mutate(class = labs[as.numeric(class)])
ggplot() +  geom_point(data = med_z, aes(x = `1`, y = `2`, color = class, alpha = 0.5)) + 
  geom_point(data = med_w, aes(x = `1`, y = `2`, fill = class), 
             shape = 25, size = 4, color = "black")  +
  facet_wrap(~time)
```

#### Latent Space Positions with MDS - see other document 

```{r, eval = F, echo = F}
wcols_t1_d1 <-grep("w\\[.,1,1", names(post))
wcols_t1_d2 <-grep("w\\[.,1,2", names(post))
wcols_t2_d1 <-grep("w\\[.,2,1", names(post))
wcols_t2_d2 <-grep("w\\[.,2,2", names(post))
wcols_t3_d1 <-grep("w\\[.,3,1", names(post))
wcols_t3_d2 <-grep("w\\[.,3,2", names(post))
zcols_d1 <- grep("z\\[.*,1\\]", names(post))
zcols_d2 <- grep("z\\[.*,2\\]", names(post))
n_post <- dim(post)[1]
post_mat <- matrix(unlist(post),nrow=n_post,byrow=F)
dist_mat1 <- parDist(x = matrix(c(post_mat[1,wcols_t1_d1],post_mat[1,zcols_d1],
                                  post_mat[1,wcols_t1_d2],post_mat[1,zcols_d2]),
                                ncol = 2), method = "euclidean") 
dist_mat2 <- parDist(x = matrix(c(post_mat[1,wcols_t2_d1],post_mat[1,zcols_d1],
                                  post_mat[1,wcols_t2_d2],post_mat[1,zcols_d2]),
                                ncol = 2), method = "euclidean") 
dist_mat3 <- parDist(x = matrix(c(post_mat[1,wcols_t3_d1],post_mat[1,zcols_d1],
                                  post_mat[1,wcols_t3_d2],post_mat[1,zcols_d2]),
                                ncol = 2), method = "euclidean") 
for (i in 2:n_post) {
  dist1 <- parDist(x = matrix(c(post_mat[i,wcols_t1_d1],post_mat[i,zcols_d1],
                                  post_mat[i,wcols_t1_d2],post_mat[i,zcols_d2]),
                                ncol = 2), method = "euclidean") 
  dist_mat1 <- dist_mat1 + dist1 
  dist2 <- parDist(x = matrix(c(post_mat[i,wcols_t2_d1],post_mat[i,zcols_d1],
                                  post_mat[i,wcols_t2_d2],post_mat[i,zcols_d2]),
                                ncol = 2), method = "euclidean") 
  dist_mat2 <- dist_mat2 + dist2
  dist3 <- parDist(x = matrix(c(post_mat[i,wcols_t3_d1],post_mat[i,zcols_d1],
                                  post_mat[i,wcols_t3_d2],post_mat[i,zcols_d2]),
                                ncol = 2), method = "euclidean") 
  dist_mat3 <- dist_mat3 + dist3
}
dist_mat1 <- dist_mat1/n_post
dist_mat2 <- dist_mat2/n_post
dist_mat3 <- dist_mat3/n_post
```

```{r, eval = F, echo = F}
mds1 <- cmdscale(dist_mat1, k=2)
mds2 <- cmdscale(dist_mat2, k=2)
mds3 <- cmdscale(dist_mat3, k=2)
#make data frame to plot
plot_mds_z <- data.frame(mds_dim1 = c(mds1[,1],mds2[,1],mds3[,1]),
                         mds_dim2 = c(mds1[,2],mds2[,2],mds3[,2]),
                         time = sort(rep(1:3,N+7)),
                         type = rep(c(rep("w",7),rep("z",N)),3)) %>%
  filter(type == "z")
plot_mds_z$class <- as.factor(c(job[,1],job[,2],job[,3]))
plot_mds_w <- data.frame(mds_dim1 = c(mds1[,1],mds2[,1],mds3[,1]),
                         mds_dim2 = c(mds1[,2],mds2[,2],mds3[,2]),
                         time = sort(rep(1:3,N+7)),
                         type = rep(c(rep("w",7),rep("z",N)),3)) %>%
  filter(type == "w")
plot_mds_w$class <- as.factor(rep(1:7,3))
ggplot() +  
  geom_point(data = plot_mds_z, aes(x = mds_dim1, y = mds_dim2, color = class, alpha = 0.5)) + 
  geom_point(data = plot_mds_w, aes(x = mds_dim1, y = mds_dim2, fill = class), 
             shape = 25, size = 4, color = "black")  +
  facet_wrap(~time)
```

#### Probability transition plots - Posterior medians 

```{r, echo = F}
med_p <- data.frame(par = rownames(fit_summ), med = fit_summ[,6]) %>%
  filter(grepl("p\\[", par)) %>%
  cSplit("par",direction = "wide") %>%
  mutate(worker = stri_extract_first_regex(par_1,"[0-9]+"),
         time = stri_extract_first_regex(par_2,"[0-9]+"),
         class = stri_extract_first_regex(par_3,"[0-9]+")) %>%
  select(-c("par_1","par_2","par_3")) %>%
  filter(time %in% c(2,3)) %>%
  mutate(past_class = c(rep(job[,2],each=7),rep(job[,3],each=7)),
         class = as.numeric(class))
trans_probs <- med_p %>% group_by(time,past_class) %>%
  mutate(tot_sum = sum(med),
         prop = med/tot_sum) %>%
  group_by(time,class,past_class) %>%
  summarise(prob_sum = sum(prop)) %>%
  mutate(past_class1 = labs[past_class],
         class1 = labs[class])
ggplot(trans_probs, aes(x = past_class1, y = class1, fill = prob_sum)) + 
  geom_tile() + facet_grid(cols = vars(time)) + 
  xlab("Past Class") + ylab("Current Class") + labs(fill='Prob. Sum') +
  ggtitle("Posterior medians for transition probabilities") + 
  theme(aspect.ratio = 1) + theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x=element_text(angle=270,hjust=1)) + 
  scale_fill_continuous(high = "#132B43", low = "#56B1F7")
```

#### Probability transition plots - Data 

```{r, echo = F}
real_trans <- data.frame(past = c(job[,1],job[,2]), current = c(job[,2],job[,3]), 
                         time = c(rep(2,N),rep(3,N))) %>%
  group_by(past,time) %>%
  mutate(tot_sum = n(),
         prop = 1/tot_sum) %>%
  group_by(past,current,time) %>%
  summarise(prob_sum = sum(prop)) %>% 
  arrange(past,time) %>%
  mutate(current1 = labs[current],
         past1 = labs[past])
ggplot(real_trans, aes(x = past1, y = current1, fill = prob_sum)) + 
  geom_tile() + facet_grid(cols = vars(time)) + 
  xlab("Past Class") + ylab("Current Class") + labs(fill='Prob. Sum') +
  ggtitle("True transition probabilities") + 
  theme(aspect.ratio = 1) + theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x=element_text(angle=270,hjust=1)) + 
  scale_fill_continuous(high = "#132B43", low = "#56B1F7")
```
