---
title: "MCMC Diagnostics - 1/23 - Simulated values as noisy initial values"
author: "Sarah Teichman"
date: "10/31/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bayesplot)
library(rstan)
library(tidyverse)
#load("~/Documents/UW/research/Model/wages_R/fit1.rda")
#load("~/Documents/UW/research/Model/wages_R/fit_sim.rda")
#load("~/Documents/UW/research/Model/wages_R/simulated_params.rda")
load("fit_test_res.rda")
```

```{r, echo = F}
fit_summ <- as.data.frame(summary(fit)$summary)
sigma_params <- sort(grep("sigma",row.names(fit_summ), value=T)[1:32])[1:16]
beta_k <- grep("beta_k",row.names(fit_summ), value=T)
u <- grep("u",row.names(fit_summ), value=T)[2:8]
v <- grep("v",row.names(fit_summ), value=T)[2:8]
q <- grep("q",row.names(fit_summ), value=T)[8:14]
other_1d <- c("alpha[1]","alpha[2]","delta[1]","delta[2]","beta[1]","beta[2]","zeta[1]","zeta[2]")
lambda <- grep("lambda",row.names(fit_summ), value=T)
eta <- grep("eta",row.names(fit_summ), value=T)
kappa <- grep("kappa",row.names(fit_summ), value=T)
```

### Overall Model Diagnostics 

Overall model diagnostics from rstan package. 

```{r, echo = F}
output_basic <- function(fit) {
  #Rhat statistics 
  print(stan_rhat(fit))
  #Ratio of effective sample size / Sample size 
  print(stan_ess(fit))
  #ratio of Monte Carlo standard error to posterior standard deviation 
  print(stan_mcse(fit))
}
output_basic(fit)
```

### Individual Parameter Diagnostics 

Individual parameter plots. Autocorrelation and trace plots for individual parameters, and histograms of posterior means for group parameters (w, z, p). 

```{r, echo = F}
get_single_plots <- function(fit, param) { 
  print(stan_ac(fit, pars = param))
  print(rstan::traceplot(fit, pars = param))
} 
get_aggreg_plots <- function(fit, param, trim = F, trim_amount) {
  ind <- grep(paste0("^",param), rownames(as.data.frame(summary(fit)$summary)))
  means <- data.frame(avg = as.data.frame(summary(fit)$summary)$mean[ind])
  print(summary(means))
  title <- paste0("Posterior Means of ",param)
  print(ggplot(means, aes(x = avg)) + geom_histogram(bins = 30) + ggtitle(title))
  if (trim == T) {
    lim <- quantile(abs(means$avg), probs = trim_amount)
    means_trim <- means %>% filter(abs(means$avg) < lim)
    print(ggplot(means_trim, aes(x = avg)) + geom_histogram(bins = 30) + 
            ggtitle(paste0(title, " Without Extremes")))
    means_out <- means %>% filter(abs(means$avg) > lim) 
    summary(means_out)
  }
}
plot_fit <- function(fit) {
  get_single_plots(fit, sigma_params)
  get_single_plots(fit, beta_k)
  get_single_plots(fit, other_1d)
  get_single_plots(fit, u)
  get_single_plots(fit, v)
  get_single_plots(fit, q)
  get_aggreg_plots(fit, "w")
  get_aggreg_plots(fit, "z")
  get_aggreg_plots(fit, "p", trim = T, trim_amount = .90)
}
plot_fit(fit)
```

```{r,echo = F}
cutoff <- 50
```

### Standardized Differences 

Histogram of $\frac{|\theta_0-\hat{\theta}|}{\theta_0}$ for all parameters, where $\hat{\theta}$ is the posterior mean. Histogram only includes standardized differences with an absolute value less than `r cutoff`. The differences excluded from the histogram are listed below. 

```{r, echo = F}
get_diff <- function(true_params, stan_ob, stat) {
  params <- unlist(true_params)
  fit_summary <- as.data.frame(summary(stan_ob)$summary)
  np <- length(params)
  if (stat == "mean") {
    post_stat <- fit_summary$mean[1:np] 
  }
  if (stat == "median") {
    post_stat <- fit_summary$`50%`[1:np]
  }
  diff <- array(data = NA, dim = np)
  for (i in 1:np) {
    diff[i] <- abs(params[i]-post_stat[i])/params[i]  
  }
  return(data.frame(diff = diff, var = names(params)))
}
diff1 <- get_diff(params, fit, "mean")
diff1_no_out <- diff1 %>%
  filter(abs(diff) < cutoff)
ggplot(data = diff1_no_out, aes(x=diff)) + geom_histogram(fill = "red", alpha = 0.5) 
per_outliers1 <- 1-nrow(diff1_no_out)/nrow(diff1)
diff1_out <- diff1 %>%
  filter(abs(diff) > cutoff) %>% 
  arrange(desc(diff))
diff1_out
```

The same as above, but with $\hat{\theta}$ as the posterior median. 

```{r, echo = F}
diff3 <- get_diff(params, fit, "median")
cutoff <- 50
diff3_no_out <- diff3 %>%
  filter(abs(diff) < cutoff)
ggplot(data = diff3_no_out, aes(x=diff)) + geom_histogram(fill = "red", alpha = 0.5, bins = 30)
per_outliers3 <- 1-nrow(diff3_no_out)/nrow(diff3)
diff3_out <- diff3 %>%
  filter(abs(diff) > cutoff) %>%
  arrange(desc(diff))
diff3_out
```